---
title: "Project 1 Code"
author: "Samikshya"
date: "2022-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(nycflights13)
library(tidyverse)
library(dbplyr)
library(ggplot2)
```
We are interested only in flights from UA so 
```{r}

data <- flights  

flights.UA <- flights  %>%
  filter(carrier == "UA") %>%
  filter(!is.na(dep_delay))
  
head(flights.UA, 10)


```

## Time of the day:  



```{r}
ggplot(data = flights.UA, aes(x = factor(hour), y = dep_delay))+
  geom_boxplot( color = 'orange')+
  xlab("hour")+
  ylab("Departure Delay")+
  ggtitle("Distribution of Departure Delay by hour")
```


```{r}
ggplot(data = flights.UA, mapping = aes(x = hour, fill = mean(dep_delay))) +
  geom_bar(color = 'blue') +
  xlab("hour")+
  ylab("count")+
  ggtitle("Median delay based on hour")
 
           
```
```

We can see that there is a pattern emerging from the boxplot. There are more variations in the box plot (Figure above) after hour 16 (i.e 4:00pm).

So, does this mean that we see more delay in the departures when it is afternoon till the .So we test: Is the difference in between departure delay between two groups (from morning hour 0 - hour 16) till afternoon to midnight (hour 16 - hour 23). 
Since our distribution of data is skewed, with longer tails towards the right, we use median to test our hypothesis. 
Null Hypothesis: 
There is no difference in mean delay between any hours in the dataset. 
Alternative: There is difference in mean departure delay based on time of hours. UA is not equal to UB. 

So, we will conduct a two tail permutation test to test the hypothesis.

Calculating the median of these two groups: 

```{r}
ggplot(data = flights.UA, aes(x = factor(hour), y = dep_delay))+
  geom_boxplot( color = 'orange')+
  xlab("hour")+
  ylab("Departure Delay")+
  ggtitle("Distribution of Departure Delay by hour")
```

```{r}

```




```{r}
ggplot(data = flights.UA, aes(x = factor(hour), y = dep_delay))+
  geom_boxplot( color = 'orange')+
  xlab("hour")+
  ylab("Departure Delay")+
  ggtitle("Distribution of Departure Delay by hour")
  
  
```

```{r}

```


```{r}
observed_hour.meandifference = mean(flights.UA$dep_delay[flights.UA$morning_afternoon == 1], na.rm = TRUE) - mean(flights.UA$dep_delay[flights.UA$afternoon_late == 1], na.rm = TRUE)
```
Observed median difference in hours for delay between morning, afternoon and is -13.39 minutes
Permutation Testing: 
The sample size for one of the group is 39024

```{r}
N <- 10^4 -1
sample.size = nrow(flights.UA)
Flighthour.size = nrow(flights.UA[flights.UA$morning_afternoon == 1,])
Flighthour.size 
result_hours.meandifference <- numeric(N)

for (i in 1:N)
{
  index = sample(sample.size, size = Flighthour.size, replace = FALSE)
  result_hours.meandifference[i] = mean(flights.UA$dep_delay[index]) - mean(flights.UA$dep_delay[-index],)
}
```


```{r}
ggplot(data=tibble(result_hours.meandifference), mapping = aes(x=result_hours.meandifference)) + 
  geom_histogram(bins= 100 ) +
  geom_vline(xintercept = observed_hour.meandifference, color = "red")+
  xlab("Result of the mean difference oberved and calculated")+
  ylab("count")+
  ggtitle("Histogram of the permutation test with observed mean difference in hour")
  
```


```{r}
p.value.hour <- ((sum(result_hours.meandifference <= observed_hour.meandifference) + 1) / (N + 1))*2
p.value.hour 
```


Conclusion: Reject the null. There is enough evidence that there is difference in the mean departure delay 

## Question 2 

Time of the Year 

```{r}
```


```{r}
```


```{r}
```


```{r}
